---
title: "Editing II: Using TidyRez"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Editing II: Using TidyRez}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
options(rmarkdown.html_vignette.check_title = FALSE)
library(rezonateR)
```

## Getting started

This tutorial is about TidyRez, the subset of functions derived from `dplyr` in Tidyverse. We will use the file saved at the end of `vignette("edit_easyEdit")`. As always, you don't have to have read that tutorial beforehand, though it may be helpful if you are new to rezonateR.

```{r}
library(rezonateR)
path = system.file("extdata", "rez007_edit2.Rdata", package = "rezonateR", mustWork = T)
rez007 = rez_load(path)
```

Unlike most other tutorials, this one will be very brief, because it assumes that you have knowledge of `dplyr`, the R package containing functions like `dplyr::mutate()` and `dplyr::select()`. If you are not familiar with `dplyr` beforehand, I suggest using a tutorial like [this](https://jhudatascience.org/tidyversecourse/wrangle-data.html) first, and then coming back to this page.

## Why TidyRez?

In general, TidyRez functions are called by adding `rez_` in front of a `dplyr` function name, such a `rez_group_by()` or `rez_mutate()`. You might wonder why you'd want to use TidyRez instead of plain `dplyr`. The main reason is that TidyRez functions allow you to keep and/or update your field access values, `inNodeMap` values, and `updateFunctions`. Using base R or classic `dplyr` functions with `rezrDF`s will result in `reload()` fails (unless you add those attributes back yourself).

Thus, TidyRez functions that add or change columns, such as `rez_mutate()` or `rez_left_join()`. will give you the option to change the field access value of that field through the `fieldaccess` field. `updateFunction`s are automatically added if you choose `auto` or `foreign`. Other TidyRez functions do not differ substantially from classic `dplyr`; they only allow you to keep your field access labels, `updateFunction`s, and `inNodeMap` values.

To see the power of TidyRez, let's try creating an emancipated `rezrDF` with only a subset of the original columns with `rez_select()`. Here, we take `trackDF$refexpr`, the table of referential expressions. We then damage one of the `auto` fields using a classic `dplyr` function. As you can see here, the emancipated `rezrDF` can still be updated using the `rezrObj`, effectively overriding the damage:

```{r}
refTable = rez007$trackDF$default %>% rez_select(id, token, chain, name, text, tokenOrderLast)
print("Before:")
head(refTable %>% select(id, tokenOrderLast))
refTable = refTable %>% mutate(tokenSeqLast = 1) #Damage refTable with a classic dplyr function
print("After:")
refTable = refTable %>% reload(rez007)
head(refTable %>% select(id, tokenOrderLast))
```

A warning is in order: TidyRez only updates *the current table*. If other tables have references to the table you're editing, they will not be updated. You must bear this in mind when using `rez_select()` and `rez_rename()`. No problems will arise if you use these functions on emancipated rezrDFs. However, if you use these functions on rezrDFs within `rezrObj`s, you should manually update any fields in other `rezrDF`s that refer to the field you've deleted or added. I plan to add a rename feature to EasyEdit in the near future that will update references from other rezrDFs.

Another implication is that if you use `rez_mutate()` and create an auto field, any references to other tables will not work. Thus, the EasyEdit function `addFieldForeign()` should be used instead. 


## What functions are available?
A few dplyr functions are completely safe to use in rezonateR, mostly those that focus on selecting rows of a table, such as `dplyr::filter()`, `dplyr::arrange()` or `dplyr::slice()`. Currently implemented TidyRez functions include:

* `rez_add_row()` for adding new entries (not recommended; `addRow()` is better)

* `rez_mutate()` for adding and editing columns

* `rez_rename()` for renaming columns

* `rez_bind_rows()` for combining rezrDFs vertically

* `rez_group_split()` for splitting rezrDFs vertically

* `rez_group_by()` and `rez_ungroup()` for grouping

* `rez_select()` for selecting certain columns inside a rezrDF

* `rez_left_join()` for left joins

Potential future additions include `rez_bind_cols()` and `rez_outer_join()`; suggestions for others are welcome if you have a use for them. The functions `rez_dfop()` and `rez_validate_fieldchange()` are used behind the scenes by TidyRez functions; if you want to create your own, please look through the documentation for these functions (and sent in a pull request when you're done with it!).

## `rez_left_join()`: A special case

Most of the TidyRez functions' syntax deviate from `dplyr` only minimally in ways that you can read about in the documentation. However, `rez_left_join()` is a notable exception.

Firstly, by default, if no suffix is specified, the suffixes are `c("", "_lower")`. That is, if you are joining two data.frames, both with a column called `name`, then the left data.frame's column will still be called `name`'` in the new data.frame but the right data.frame's column will get called `name_lower`. Because `_lower` is not a very informative name, it's best to supply your own suffixes.

In addition to a `fieldaccess` field, as we've mentioned before, you will also need:

*  `rezrObj` which is self-explanatory
* `fkey` - the name of the field in the first rezrDF that corresponds to IDs of the second rezrDF. If this is not specified, it will be guessed from the `by` argument of `dplyr::left_join()`.
* `df2key` - the name of the field in the second rezrDF that corresponds to `fkey`. If this is not specified, it will be guessed from the `by` argument of `dplyr::left_join()`.
* An `df2Address` field a string that tells `rez_left_join()` how to find the source rezrDF from the rezrObj next time.

Addresses are mostly used by `rezonateR` under the hood, but in the case of `rez_left_join()`, you do need to be able to use it. If the source `rezrDF` doesn't belong to a layer, e.g. `tokenDF`, then the DF name is the address. If the source `rezrDF` belongs to a layer, put a '/' between the table and the layer, e.g. `'trackDF/refexpr'`. Don't forget that default layers are also layers!

As an example, we'll take `averageWordLength` in the `unitDF`, which we added in `vignette("edit_easyEdit")`, and add it back to `tokenDF`, so that when we look at each word, we know the average length of the word in its units. Here's how:

```{r}
rez007$tokenDF = rez007$tokenDF %>% rez_left_join(
  y = rez007$unitDF %>% rez_select(id, averageWordLength),
  by = c("unit" = "id"),
  fieldaccess = "foreign",
  df2Address = "unitDF",
  fkey = "unit",
  rezrObj = rez007
)
rez007$tokenDF %>% select(id, text, averageWordLength) %>% head
```


## Onwards!
Using EasyEdit or TidyRez, it is not hard to use some rules to add some automatic annotations, and then correct them by hand in a spreadsheet program. The next tutorial, `vignette("edit_external")`, will cover exactly this use case. We will export data to a `.csv`, edit it outside R, and then import it back.

As always, saving is a virtue!

```{r}
savePath = "rez007.Rdata"
rez_save(rez007, savePath)
```
