---
title: "Tracking down coreference phenomena"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tracking down coreference phenomena}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
options(rmarkdown.html_vignette.check_title = FALSE)
library(rezonateR)
```

## Getting started

This tutorial discusses trees in `rezonateR`. In the first tutorial `vignette("import_save_basics")`, we took a quick glimpse at trees when discussing the node map, but now it's time to dive a bit deeper! And if you haven't read the previous vignettes, as always, that's fine because each tutorial is self-contained. I won't assume you've read the bit about trees in `vignette("import_save_basics")`.


We will be using the same Santa Barbara Corpus as before:

```{r}
library(rezonateR)
path = system.file("extdata", "rez007_track.Rdata", package = "rezonateR", mustWork = T)
rez007 = rez_load(path)
```

The file contains predicate-argument structures for the first fifth of the text or so. In practice, trees are often the most time-consuming data structure to process if you have annotated them throughout the whole text. Each tree is a predicate-argument structure with a tree as the root and all its arguments as leaves. In cases of IUs with more than one verb, multiple trees are created, one for each verb.

## Speaking of tracks ...

As usual, let's not forget:

```{r}
savePath = "rez007.Rdata"
rez_save(rez007, savePath)
```


